---
title: "EFA"
output: html_notebook
---

```{r}
library(tidyverse)
library(psych)
library(GPArotation)
```

```{r}
df = read.csv(readline("Path of dataframe (e.g., C:/experiment/df.csv): "))
```

# 1. EFA with 100<= sample size

## 1.1. Statistical learning measures
```{r}
SL_100plus_cor = df %>% 
  select(
    SEGM_AL_medRT_train, SEGM_AL_medRT_TRN3_RND4, SEGM_AL_medRT_RND4_REC5,
    SEGM_AL_ACC_train, SEGM_AL_ACC_TRN3_RND4, SEGM_AL_ACC_RND4_REC5,
    SEGM_AL_2AFC_bigram, SEGM_AL_2AFC_trigram, SEGM_AL_SEGM_prod_data,
    AGL_ACC_train, AGL_2AFC_phr, AGL_2AFC_sent, AGL_prod
  ) %>% 
  cor(use = "pairwise.complete.obs")
SL_100plus_cor %>% 
  scree(factors = FALSE)
```

```{r}
SL_EFA_100plus = df %>% 
  select(
    SEGM_AL_medRT_train, SEGM_AL_medRT_TRN3_RND4, SEGM_AL_medRT_RND4_REC5,
    SEGM_AL_ACC_train, SEGM_AL_ACC_TRN3_RND4, SEGM_AL_ACC_RND4_REC5,
    SEGM_AL_2AFC_bigram, SEGM_AL_2AFC_trigram, SEGM_AL_SEGM_prod_data,
    AGL_ACC_train, AGL_2AFC_phr, AGL_2AFC_sent, AGL_prod
  ) %>% 
  fa(nfactors = 4)
```
